\chapter{Remarks for AKP}

\begin{enumerate}
\item Eventially, everywhere lower index in the arrays of points (and other things) will be changed to upper index.
The reasons 
(1) coordinates in Riemannian geometry usually come with upper idexes and it is more natural to write $x^i=\dist{a^i}{}{}$ 
(2) some times we need sequences of arrays, then it is better to write $a^i_n$ than $a_{i,n}$.
There is a little problem once you want to use prime --- I would suggest to use $\acute{a}$ instead of $a'$ in such cases.

\item I plan to change the notation of GH-convergence from $X_n\GHto X$ to $X_n\xto{\GH} X$;
here $\GH$ stays for the topology on $\bigsqcup X_i$ which defines Gromov--Hausdorff convergence. 
The main reason is that for induced convergences we use $\to$, or $\rightrightarrows$ $\rightharpoonup$ and it looks very unnatural. 

\item I removed $\backslash$olim and $\backslash$oto everywhere. Instead we write $\lim_{n\to\o}$ or $x_n\to x_\o$ as $n\to\o$.

\item DONE: $\backslash$underline $\to$ $\backslash$ushort

\item DONE: change ``induced'' to ``inherited/intrinsic''

\item $T$ to $\ell$.

\item At the moment $\proj$ ($\backslash$proj) and $\varpi{}$ ($\backslash$varpi) are nearly identical...

\item $\sup_{x\in X}\{f(x)\}$ to $\sup\set{f(x)}{x\in X}$;
the same for $\inf$, $\max$, $\min$ and so on.


\item extremal subset $\to$ extremal set

\item remove ``native''

\item mid-points OR midpoints?  

S: midpoints.

\item We seem to agree to use terms
``geodesic'' mean minimizing geodesic;
``local geodesic'' means geodesic;
``unit-speed geodesic'' and the term geodesic can be also used for
 the image of minimizing geodesic
the class of all constant speed reparametrizations of a geodesic 

\item I think it is better to change everywhere $\ZZ_{\ge0}$ to $\ZZ_+$ and $\RR_{\ge0}$ to $\RR_+$ it suits better with $\SS_+$ (and $\SS_{\ge0}$ does not have sense)

\item Should we change to Bourbaki-notation of real intervals? As $\l]a,b\r[$ instead of $(a,b)$?

\item We use $\partial$ for the set of supporting vectors, for the boundary, and S. wants to use it for frontier(=rel. boundary). We also use $\Int$ for interior --- it seems more natural to use some letters for frontier...

\item Stephanie suggests to change $\Fr$ to $\partial$; in positve curvature might be  a bit of confusion between rel. boundary and boundary. We might use $\partial_\spc{X}A$ for rel boundary of $A\subset \spc{X}$ and $\partial\spc{X}$ for boundary. I also use $\partial_p f$ for subgradient...

\item in the end we need to change $\l]x^1x^2\r[$ to something closer to 
$\,{]}x^1x^2{[}\,$.

\item DONE: $G_\delta$ $\to$ G-delta

\item DONE: Some people write $\oBall(p,R)$ and $\cBall[p,R]$. 
Maybe it is better this way --- ``$)$'' reminds open interval and ``$]$'' closed...

\item in the end check usage of CBA and CBB...

\item me might write $(1{+}3)\mc\kappa$-point comparison or point-on-side${}\mc\kappa$ comparison???

S: I think we need   ``$\kappa$-comparison''.

\item (2+2)-comparison $\to$ (2+2)-point comparison.

\item We might want to rename ultratangent cone to ultrablowup

\item I want to rename the general construction 
in baricentric simplexex to something else
and use the name "$\kappa$-barycentric simplex"
for only for arays modified distance functions
($\md\kappa\circ\dist{x^i}{}{}$).

I think of a good name for general construction,
but can not come up with anything better than
"argmin map".

Any suggestions?

S: "minpoint map".

\item $\not\in$ and others split wrong at the end of the line...

\item I think to change expressions like $|x-y|<\eps$ to $x\approx y\pm \eps$.
(I noticed that the notation $x=y\pm \eps$ is used for $x=y$ with absolute error $\eps$,
but we use ``$\pm$'' for ``$+$'' ar ``$-$'', so ``$\approx$'' instead of ``$=$'' will remind that meaning is different).
\subitem I would also use $\approx$ with little $o$ notation.

\subitem A. hates $o(x)$ notation.
We need a notation for $o(x)/x$. 
One candidate is $\dot o_x$, so for $o(x\cdot\ln x)$ we could write $\dot o_x\cdot x\cdot\ln x$ 
Perelman use $\varkappa(x)$.
Can you suggest anything better?


\subitem A possible solution is to use $\lessgtr$ or $\leftrightharpoons$for example
\[x\lessgtr y\pm \eps(t)\]
\[x\leftrightharpoons y\pm \eps(t)\]
\item use term ``linearly differentiable'' if differential is defined and linear.

\item $\vol^m$ or $\vol_m$?

\item $\imath$ $\to$ $\iota$?

\item change supporting vector to maybe subnormal?

\item subergraph $\to$ epigraph; subgraph $\to$ hypograph? 

\item in line use either $\tfrac{\varpi\kappa}{2}$ $\to$ $\varpi\kappa/2$

\item what is better?
\[\mangle\hinge p q x
=
\lim_{
\substack{
\bar x\to p
\\
\bar x\in\,\l]px\r]}}\angk\kappa p q{\bar x}\]
or 
\[\mangle\hinge p q x
=
\lim_{\bar x\to p}
\set{\angk\kappa p q{\bar x}}{\bar x\in\,\l]px\r]}\]

\item polygon line OR broken geodesic

\item $\eps$-Hausdorff approximation OR $\eps$-isometry --- BBI use  $\eps$-isometry.

\item metion authors around  cites 

\item unitspeed $\to$ unit-speed and constantspeed $\to$ constant-speed?

\item do we need pregeodesic?

\item better name for $\Str$?

\item Fix pagebrake before $\backslash$qedsf and $\backslash$claimqedsf

\item Define distance maps and their liftings

\item What if we change Gromov--Hausdorff to Hausdorff?
By the way, Hausdorff defined so called closed and open convergences.
They are better than Hausdorff convergence, but the later became mor popular by stupid reason.
It also made Gromov define convergence wrong way.

\item vertexes $\to$ verteces

\item subnormal $\to$ normal to supporting hyperplane

\item sequence is a ``lifting'' or ``liftings''?

\item CHECK Lipschitz/locally Lipschitz and co-Lipschitz/ locally co-Lipschitz.

\item unify ``if'' and ``only if'' parts.

\item good name for radial curves --- base point, curvature, and starting point???

\item Remove new par after claim

\item $A\times_f B$ to $A\mathrel{{\times}_f} B$.

\item Tangent space to tangent cone --- we use tangent $\kappa$-cone sometimes.

\item semi-continuous semi-concave OR semicontinuous semiconcave.

\item N.B. or Pay attention?

\item i.e. or that is it?

\item Reshetnyak's puth pastry $\to$ Reshetnyak's building

\item $[xy]$ $\to$ $[x,y]$.

\item Chauchy sequence $\to$ sequence conveges in itself

\item What about $\CAT{-\infty}$ spaces?

\item Do we need $R$-length spaces?

\item Eventually we will need to change notation for the subfunctions.
They are essentially continous partially defined functions.
I would say ``function'' for ``partially defined functions''
and ``total function'' for ``function''.

I propose the following system of notations:

$f\:X\to Y$ for partially defined function

$f\:X\shortmid\to Y$ for function       

$f\:X\shortmid\to\shortmid Y$     for onto functon

$f\:X\leftrightarrow Y$      for injective partially defined function

$f\:X\shortmid\leftrightarrow\shortmid Y$     for bijection

and so on.

\end{enumerate}

Notations:
\begin{enumerate}
\item $p'$ usually denotes $p'\approx p$; 
and $\acute{p}^i\approx p^i$
\item $\bar x$ denotes a point on  $]p x]$ (for an evident choice of $p$)
\item The origin of tangent cone $\T_p$ is denoted by $o$ or $o_p$. BUT it would be more intuiteve to denote it by something like $\vec p$ or $\dot p$ --- you may also write $\ddir pp$, but that is really stupid.
It would be also good  to use the same modifier for shortcut like $\dot q=\log p q$
\item $\backslash$map and $\backslash$map[1] gives $\map$, $\backslash$map[2] gives $\map[2]$, $\backslash$map[3] gives $\map[3]$.
\item $\oBall(S,R)$  to denote an $R$-neighborhood of a set $S$.
\end{enumerate}

